### 2015-06-26 ids-fusion.tex
Flavors of IDS --- kicked in a sentence to clarify

Existing \idses are not designed to work together as part of a
suite of sensors.  Instead, each program generates a separate,
often voluminous, stream of reports; fusing them into a coherent 
overview is left
% TODO restore to QR15
%as an exercise for
 to
the user. 
Ideally, network administrators would have a suite of different \idses active,
since different \ids approaches
% TODO restore to QR15
% --- network (NIDS) vs.\
% host-based (HIDS), signature vs.\ anomaly-based --- 
bring different
strengths and weaknesses.
%%
%% Hm, can we just get rid of the details NIDS/HIDS and
%% signature/anomaly comparisons, and keep just the takeaway? I don't
%% believe we refer to them after this. 
%%
Briefly, more accurate \idses have significantly higher software
administration burdens, and there is an inevitable tradeoff between
false positives and false negatives.
%%
%%Some systems aim to overcome the limitations of existing \idses by assembling 
%%a suite of
%%sensors.  This can be a very efficient way to overcome the problem
%%of false positives, but relies crucially on finding sensors that fail
%%relatively independently.  
%%
%% 
%% NIDS systems are easy to deploy; from only one
%% or two locations, they can watch all relevant traffic.  Unfortunately, they also have
%% substantial problems with false positives and false negatives, because they
%% don't ``know'' the meaning of the traffic they see.  Often they must draw
%% conclusions based only on information in packet headers (although some
%% systems do attempt to reconstruct higher level protocols from packet data),
%% and more extensive reasoning about traffic either would slow down the
%% network, or must ignore some packets altogether.
%% A HIDS may take into account much more information about the state
%% of a defended system, but installing HIDS pervasively involves deploying and
%% configuring much more software.  Typically HIDS are installed on only the most
%% critical assets.
%% 
%% Signature-based systems attempt to match patterns of known bad behavior.  Such
%% \idses are plagued by false negatives, both because of novel attacks for which
%% there are no patterns to match, and because of known %knowledgeable attackers take
%% countermeasures against signature identification (e.g.\ encrypting malware, or
%% permuting its structure).  Anomaly detectors are prone to false
%% positives: anomalies are often benignly anomalous,
%% not maliciously so.  Detectors often believe there is an
%% anomaly when none exists because computer usage patterns are
%% notoriously difficult to learn: they are often non-stationary, they vary
%% substantially from location-to-location, etc.
%% \draft{Add citations of papers on the challenges of anomaly detection.}

>>>>

Existing \idses are not designed to work together as part of a
suite of sensors.  Instead, each program generates a separate,
often voluminous, stream of reports; fusing them into a coherent 
overview is left
% TODO restore to QR15
%as an exercise for
 to
the user. 
Ideally, network administrators would have a suite of different \idses active,
since different \ids approaches
% TODO restore to QR15
% --- network (NIDS) vs.\
% host-based (HIDS), signature vs.\ anomaly-based --- 
bring different
strengths and weaknesses.
%%
%% Hm, can we just get rid of the details NIDS/HIDS and
%% signature/anomaly comparisons, and keep just the takeaway? I don't
%% believe we refer to them after this. 
%%
Briefly, more accurate \idses have significantly higher software
administration burdens, and there is an inevitable tradeoff between
false positives and false negatives.
This tradeoff applies for all of the different varieties of IDS,
for example NIDS or HIDS, signature-based, anomaly-detecting, various
hybrids.  We do not discuss the varieties of \idses here;
we are not concerned with the mechanics of clustering reports, but
rather with the subsequent assessment of hypothesized events.
%%
%%Some systems aim to overcome the limitations of existing \idses by assembling 
%%a suite of
%%sensors.  This can be a very efficient way to overcome the problem
%%of false positives, but relies crucially on finding sensors that fail
%%relatively independently.  
%%
%% 
%% NIDS systems are easy to deploy; from only one
%% or two locations, they can watch all relevant traffic.  Unfortunately, they also have
%% substantial problems with false positives and false negatives, because they
%% don't ``know'' the meaning of the traffic they see.  Often they must draw
%% conclusions based only on information in packet headers (although some
%% systems do attempt to reconstruct higher level protocols from packet data),
%% and more extensive reasoning about traffic either would slow down the
%% network, or must ignore some packets altogether.
%% A HIDS may take into account much more information about the state
%% of a defended system, but installing HIDS pervasively involves deploying and
%% configuring much more software.  Typically HIDS are installed on only the most
%% critical assets.
%% 
%% Signature-based systems attempt to match patterns of known bad behavior.  Such
%% \idses are plagued by false negatives, both because of novel attacks for which
%% there are no patterns to match, and because of known %knowledgeable attackers take
%% countermeasures against signature identification (e.g.\ encrypting malware, or
%% permuting its structure).  Anomaly detectors are prone to false
%% positives: anomalies are often benignly anomalous,
%% not maliciously so.  Detectors often believe there is an
%% anomaly when none exists because computer usage patterns are
%% notoriously difficult to learn: they are often non-stationary, they vary
%% substantially from location-to-location, etc.
%% \draft{Add citations of papers on the challenges of anomaly detection.}
